### 2022-11-28 draft.tex
Deploying to gh-pages from @ graciellehigino/ms_range_interactions@281f7033735c759bbd6895f8b283e9e1e1155977 🚀

Finding a species in a certain location is like finding an encrypted
message that traveled through time. It carries the species' evolutionary
history, migration patterns, as well as any direct and indirect effects
generated by other species (some of which we may not even know exist).
Ecologists have been trying to decode this message with progressively
more powerful tools, from their field notes to highly complex
computational algorithms. However, to succeed in this challenge it is
important to have the right clues in hand. There are many ways we can be
misled by data - or the lack of it: taxonomic errors (e.g., due to
updates in the taxonomy of a species), geographic inaccuracy (e.g.,
approximate coordinates or lack of documentation about their accuracy),
or sampling biases (e.g.~data clustered near roads or research centers)
(Ladle and Hortal 2013; Hortal et al. 2015; Poisot et al. 2021). One way
to identify - and potentially fix - these errors is to combine many
different pieces of information about the occurrence of a species, so
agreements and mismatches can emerge. Although previous studies have
combined different types of occurrence data to measure the accuracy of
datasets (Hurlbert and Jetz 2007; Hurlbert and White 2005; Ficetola et
al. 2014), none have used different types of information so far (i.e.,
ecological characteristics other than geographical distribution). Here
we suggest jointly analysing species occurrence (range maps and point
occurrences) and ecological interactions to identify mismatches between
datasets.

@@

A significant challenge in this debate is the quality and quantity of
species distribution and ecological data (Boakes et al. 2010; Ronquillo
et al. 2020; Meyer, Weigelt, and Kreft 2016) - a gap that can lead to
erroneous conclusions in macroecological research (Hortal et al. 2008).
Amongst the geographical data available are the range maps provided by
the International Union for the Conservation of Nature (IUCN). Such maps
consist of simplified polygons, often created as alpha or convex hulls
around known species locations, refined by expert knowledge about the
species (IUCN Red List Technical Working Group 2019). These maps can be
used in macroecological inferences in the lack of more precise
information (Fourcade 2016; Alhajeri and Fourcade 2019), but it has been
recommended that they are used with caution since they tend to
underestimate the distribution of species that are not well-known
(Herkt, Skidmore, and Fahr 2017) (especially at fine scale resolutions;
Hurlbert and Jetz (2007); Hurlbert and White (2005)), do not represent
spatial variation in species occurrence and abundance (Dallas, Pironon,
and Santini 2020), and can include inadequate areas within the estimated
range. Another source of species distribution information is the Global
Biodiversity Information Facility (GBIF), which is an online repository
of georeferenced observational records that come from various sources,
including community science programs, museum collections, and long-term
monitoring schemes. A great source of bias in these datasets is the
irregular sampling effort, with more occurrences originating from
attractive and accessible areas and observation of charismatic species
(Alhajeri and Fourcade 2019). As for ecological data, a complete
assessment is difficult and is aggravated by biased sampling methods and
data aggregation (Poisot et al. 2020; Hortal et al. 2015). Nevertheless,
we have witnessed an increase in the availability of biodiversity data
in the last decades, including those collected through community science
projects (Callaghan et al. 2019; Pocock et al. 2015) and dedicated
databases, such as Mangal (Poisot et al. 2016). This provides an
opportunity to merge species distribution and ecological interaction
data to improve our predictions of where a species may be found across
large spatial scales (e.g., continental and global).

It has been demonstrated that the agreement between range maps and point
data varies geographically (Hurlbert and Jetz 2007; Hurlbert and White
2005; Ficetola et al. 2014). Adding ecological interaction data to this
comparison might help to elucidate where these (dis)agreements are more
likely to be true and which dataset better represent the actual
distribution of species. In this context, we elaborate a method that
allows us to detect areas of potential misestimation of species'
distribution data (more precisely range maps) based on interaction data.
This method is based on the assumption that organisms cannot persist in
an area unless they are directly or indirectly connected to a primary
producer within their associated food web (Power 1992). Thus, given that
herbivores are the main connection between plant resources (directly
limited by environmental conditions) and predators (Dobson 2009; Scott
et al. 2018), the range of a predator (omnivore or carnivore) depends on
the overlapping ranges of its herbivore preys. If sections of a
predator's range does not overlap with at least one of its prey it will
become disconnected from primary producers, and therefore we would not
expect the predator to occur in this area.

This mismatch can be the result of different mechanisms, like the
misestimation of both the predator's and the preys' ranges (Ladle and
Hortal 2013; Rondinini et al. 2006), taxonomic errors (Isaac, Mallet,
and Mace 2004; Ladle and Hortal 2013), or the lack of information about
trophic links (i.e., the lack of connection between the ranges of a
predator and a primary producer may be due a third species we don't know
is connected to both). Here we discuss the mechanisms that can lead to
the lack of agreement between data, and build from that a vision for the
next steps, reinforcing the importance of geographically explicit
interaction data.

@@

We indentified areas of uncertainty on the ranges of predators based on
a simple rule: we removed any part of a predator's range that did not
intersect with the range of at least one prey herbivore species, which
in turn is directly connected to a primary producer (plants). To do
that, we used a Serengeti food web dataset (Baskerville et al. 2011)
(which comprises carnivores, herbivores, and plants from Tanzania) and
its species ranges from IUCN. Then, we calculated the difference in
range sizes between the original IUCN ranges of predators and those
without the areas where they would be alone, based on species
interaction data. Finally, we added the GBIF occurrence points for the
Serengeti species to investigate whether the results would be different
if we used another source of distribution data. \#\# Data

@@

For each species in the dataset we collated point observation data from
GBIF (www.gbif.org). We used the GBIF download API to retrieve all
species occurrences on November 22nd 2022 (GBIF.org 2022). We restricted
our query to the data with spatial coordinates and which were inside the
spatial extent of our rasters. We did not use continental or date
filters to retrieve as much data as possible. However, a few
observations were localized in the ocean near latitude 0° and longitude
0°. We assumed these were errors and removed all observations falling in
the extent between latitudes 2°S and 2°N and longitudes 2°W and 2°E to
keep only mainland sites.

@@

Mammal species found in the Serengeti food web are widespread in Africa,
especially in grasslands and savannahs (first panel of
fig.~\ref{fig:richness}). However, most local networks (83.2\%) built
using the original IUCN range maps had at least one mammal species
without a path to a primary producer (second panel of
fig.~\ref{fig:richness}). On average, local food webs had almost the
third of their mammal species (mean = 30.5\%, median = 14.3\%)
disconnected from basal species. In addition, many networks (16.6\%)
only had disconnected mammals; these networks however all had a very low
number of mammal species, specifically between 1 and 4 (from a total of
32). As expected, the proportion of carnivores with a path to a primary
producer was conditional on the total number of mammal species in each
local network (third panel of fig.~\ref{fig:richness}).

@@

\hypertarget{specialized-predators-lose-more-range}{%
\subsection{Specialized predators lose more
range}\label{specialized-predators-lose-more-range}}

\begin{figure}
\hypertarget{fig:degree}{%
\centering
\includegraphics{figures/rel_loss-outdegree-species.png}
\caption{Negative relationship between the out degree of predator
species and their relative range loss. More specialized predators lose a
higher proportion of their ranges due to mismatches with the ranges of
their preys.}\label{fig:degree}
}
\end{figure}

Predators with fewer prey lose more range with our method
(fig.~\ref{fig:degree}). For instance, both \emph{Leptailurus serval}
and \emph{Canis mesomelas} have only one prey in the Serengeti food web
(tbl.~\ref{tbl:everyone}), each of them with a very small range compared
to those of their predator. This discrepancy between range sizes
promotes significant range loss. On the other hand, predators of the
genus \emph{Panthera} are some of the most connected species, and they
also lose the least proportion of their ranges. This mismatch between
predators and preys can also be a result of taxonomic disagreement
between the geographical and ecological data. Although \emph{Canis
aureus} has the same number of prey as \emph{Caracal caracal}, none of
the prey taxa of the former occurs inside its original range
(tbl.~\ref{tbl:everyone}), which results in complete range loss.

@@

The proportion of GBIF pixels (pixels with at least one GBIF occurrence)
falling within the IUCN ranges varied from low to high depending on the
species (fig.~\ref{fig:gbif}, left). The lowest proportions occurred for
species with small ranges. Amongst herbivores, \emph{Rhabdomys pumilio}
has a proportion of 22.6\% of its presence pixels within its IUCN range,
while predators have this proportion above 55\% (such as \emph{Lycaon
pictus}, with 55.1\%, and \emph{Canis aureus}, with 56.2\%).
Nevertheless, some species with smaller ranges showed high data overlap
(such as \emph{Canis mesomelas}, with 94.9\%, and many herbivores).
Overall, predators and preys displayed similar overlap variations, and
species with median and large ranges had higher proportions of
occurrences falling into their IUCN range.

The proportion of GBIF pixels in updated ranges can only be equal to or
lower than that of the original ranges, as our analysis removes pixels
from the original range and does not add new ones. Rather, the absence
of a difference between the two types of ranges indicates that no pixels
with GBIF observations, hence likely true habitats, were removed by our
analysis. Here this proportion was mostly similar to that of the
original IUCN ranges for most predator species (fig.~\ref{fig:gbif}).
Two species showed no difference in proportion while four species showed
only small differences (\emph{Crocuta crocuta} lost 1.3\% of the
original data overlap; \emph{Acinonyx jubatus} lost 1.9\%;
\emph{Panthera pardus} lost 8.8\%; and \emph{Caracal caracal} lost
12.3\%). On the other hand, three species, \emph{Canis aureus},
\emph{Canis mesomelas}, and \emph{Leptailurus serval} showed very high
differences, with overlaps lowered by 100\%, 57.4\%, and 100\%
respectively. These last two species are also the only predators with a
single prey in our metaweb. \emph{Canis aureus} has four preys, but it
has one of the smallest ranges in IUCN, which is not covered by any of
its preys. This result reinforces the concern raised in the literature
on the use of IUCN range maps for species that are not well known
(Herkt, Skidmore, and Fahr 2017), demonstrating how small range species
are likely to have their distribution underestimated in the IUCN
database. Additionally, the fact that \emph{Canis aureus} had such a
conspicuous mismatch between both the original and updated IUCN range
maps, and between GBIF and IUCN data, may indicate a taxonomic
incongruency between the three databases used here, which we explore in
the Discussion section. Our results delineate how a mismatch between
GBIF and IUCN databases differ greatly with small changes in herbivore
species ranges, and it is somewhat positively related to range size for
predator species. Moreover, we show that accounting for interactions
does not necessarily aggravates this dissimilarity, but it is relevant
for species with little ecological information or specialists.

\begin{figure}
\hypertarget{fig:gbif}{%
\centering
\includegraphics{figures/gbif_panels.png}
\caption{Left panel: Distribution of the proportion of GBIF pixels
(pixels with at least one occurrence in GBIF) falling into the IUCN
range for different range sizes. Right panel: Differences between the
proportion of GBIF pixels falling into the IUCN and the updated ranges
for every predator species. Arrows go from the proportion inside the
original range to the proportion inside the updated range, which can
only be equal or lower. Overlapping markers indicate no difference
between the types of layers. Species markers are the same on both
figures, with predators presented in distinct colored markers and all
herbivores grouped in a single grey marker. Pixels represent a
resolution of 10 arc-minutes.}\label{fig:gbif}
}
\end{figure}

@@

The case of the golden jackal (\emph{Canis aureus}) is a good
illustration of how the taxonomic, geographical and ecological data can
be used to validate one another. The jackal is a widespread taxon in
northern Africa, Europe, and Australasia, generally well adapted to
local conditions due to its largely varied diet (Tsunoda and Saito 2020;
Krofel et al. 2021). Because of that, we expected that
the~\emph{Canis}~species in our dataset would be the ones losing the
least amount of range, with a higher value of the proportion of GBIF
pixels within their IUCN range maps. However, the taxonomy of this group
is a matter of intense discussion, as molecular and morphological data
seem to disagree in the clustering of species and subspecies (Krofel et
al. 2021; Stoyanov 2020). This debate probably influenced our results:
with originally only 56.2\% of the GBIF pixels of the golden jackal
overlapping with its IUCN data, we suspect that many of the GBIF
occurrences refer to other~\emph{Canis}~species, and that its taxonomic
identification in the network database is probably outdated. This led to
a complete exclusion of~\emph{Canis aureus}~from its original range in
our analysis, despite the fact that this species has four documented
preys in our metaweb.

@@

First, \emph{Panthera leo} was one of the species with no difference
between ranges before and after our analysis, but 40.7\% of the GBIF
occurrences for this species fell outside its IUCN range
(fig.~\ref{fig:gbif}). In this particular case, the IUCN maps seem to
agree with species interaction data. However, the disagreement between
the IUCN and the GBIF databases is concerning and suggests that the IUCN
maps might underestimate the lion's distribution.

On the other hand,~\emph{Leptailurus serval}~and~\emph{Canis
mesomelas}~are two of the three species that have the higher proportion
of mismatched range due to the lack of paths to a herbivore, but are
also some of the species with the higher proportion of GBIF occurrences
inside IUCN range maps (fig.~\ref{fig:gbif}). This indicates that the
information we are missing for these two species is related to either an
additional interaction or to the presence of interacting species. To
illustrate that, we mapped the GBIF data for the prey
of~\emph{Leptailurus serval}, with a mobility buffer around each point
(fig.~\ref{fig:serval}). When considering GBIF data, approximately 42\%
of the prey's occurrences are within the portion of the predator's range
that was divergent from its original IUCN data. With the buffer area,
this corresponds to 15\% of the mismatched area. By adding GBIF
information for the prey, we could therefore reduce the discrepancy of
the range (or information) for the predator by 15\% since its
distribution is conditional on the occurrence of its preys. In other
words, the range mismatch was exagerated because we were missing
information on the presence of an interacting species (i.e., this also
indicates that there is a mismatch - or complementarity - between the
IUCN and GBIF data for their prey).

@@

Here we demonstrated how we can detect uncertainty in species
distribution data using ecological interactions. Knowing where
questionable occurrence data are can be crucial in ecological modelling
(Hortal 2008; Ladle and Hortal 2013), and accounting for these errors
can improve model outputs by diminishing the error propagation (Draper
1995). For instance, we believe this is a way to account for ecological
interactions in habitat suitability models without making the models
more complex, but by making sure (not assuming) that the input data -
the species occurrence - actually accounts for ecological interactions.
It is important to notice, however, that the quality and usefulness of
this method are highly correlated with the amount and quality of data
available about species' occurrences \textbf{and} interactions. With
this paper, we hope to add to the collective effort to decode the
encrypted message that is the occurrence of a species in space and time.
A promising avenue that adds to our method is the prediction of networks
and interactions in large scales (Strydom et al. 2021; Windsor et al.
2022), for they can add valuable information about ecological
interactions where they are missing. Additionally, in order to achieve a
robust modelling framework towards actual species distribution models we
should invest in efforts to collect and combine open data on species
occurrence and interactions (Windsor et al. 2022), especially because we
may be losing ecological interactions at least as fast as we are losing
species (Valiente-Banuet et al. 2015).

@@

We acknowledge that this study was conducted on land within the
traditional unceded territory of the Saint Lawrence Iroquoian,
Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak nations. We
thank the editor and reviewer for their thoughtful comments, which
considerably improved this manuscript.

>>>>

Finding a species in a certain location is like finding an encrypted
message that traveled through time. It carries the species' evolutionary
history, migration patterns, as well as any direct and indirect effects
generated by other species (some of which we may not even know exist).
Ecologists have been trying to decode this message with progressively
more powerful tools, from their field notes to highly complex
computational algorithms. However, to succeed in this challenge it is
important to have the right clues in hand. There are many ways we can be
misled by data - or the lack of it: taxonomic errors (e.g., due to
updates in the taxonomy of a species), geographic inaccuracy (e.g.,
approximate coordinates or lack of documentation about their accuracy),
or sampling biases (e.g.~data clustered near roads or research centers)
(Ladle and Hortal 2013; Hortal et al. 2015; Poisot et al. 2021). One way
to identify - and potentially fix - these errors is to combine many
different pieces of information about the occurrence of a species, so
agreements and mismatches can emerge. Although previous studies have
combined different types of occurrence data to measure the accuracy of
datasets (Hurlbert and Jetz 2007; Hurlbert and White 2005; Ficetola et
al. 2014), none have used different types of information so far (i.e.,
ecological characteristics other than geographical distribution). Here
we suggest jointly analysing species occurrence (range maps and point
occurrences) and ecological interactions to identify mismatches between
datasets and areas of data deficit.

@@

A significant challenge in this debate is the quality and quantity of
species distribution and ecological data (Boakes et al. 2010; Ronquillo
et al. 2020; Meyer, Weigelt, and Kreft 2016) - a gap that can lead to
erroneous conclusions in macroecological research (Hortal et al. 2008).
Amongst the geographical data available are the range maps provided by
the International Union for the Conservation of Nature (IUCN). Such maps
consist of simplified polygons, often created as alpha or convex hulls
around known species locations, refined by expert knowledge about the
species (IUCN Red List Technical Working Group 2019). These maps can be
used in macroecological inferences in the lack of more precise
information (Fourcade 2016; Alhajeri and Fourcade 2019), but it has been
recommended that they are used with caution since they tend to
underestimate the distribution of species that are not well-known
(Herkt, Skidmore, and Fahr 2017) (especially at fine scale resolutions;
Hurlbert and Jetz (2007); Hurlbert and White (2005)), do not represent
spatial variation in species occurrence and abundance (Dallas, Pironon,
and Santini 2020), and can include inadequate areas within the estimated
range. Another source of species distribution information is the Global
Biodiversity Information Facility (GBIF), which is an online repository
of georeferenced observational records that come from various sources,
including community science programs, museum collections, and long-term
monitoring schemes. A great source of bias in these datasets is the
irregular sampling effort, with more occurrences originating from
attractive and accessible areas and observation of charismatic species
(Alhajeri and Fourcade 2019). As for ecological data, a complete
assessment is difficult and is aggravated by biased sampling methods,
data aggregation (Poisot et al. 2020; Hortal et al. 2015) and by the
fact that interactions are very often events that occur in a narrow
window of time. Nevertheless, we have witnessed an increase in the
availability of biodiversity data in the last decades, including those
collected through community science projects (Callaghan et al. 2019;
Pocock et al. 2015) and dedicated databases, such as Mangal (Poisot et
al. 2016). This provides an opportunity to merge species distribution
and ecological interaction data to improve our predictions of where a
species may be found across large spatial scales.

It has been demonstrated that the agreement between range maps and point
data varies geographically (Hurlbert and Jetz 2007; Hurlbert and White
2005; Ficetola et al. 2014). Adding ecological interaction data to this
comparison might help to elucidate where these (dis)agreements are more
likely to be true and which dataset better represent the actual
distribution of a species. In this context, we elaborate a method that
allows us to detect areas of potential misestimation of species'
distribution data (more precisely range maps) based on interaction data.
This method is based on the assumption that organisms cannot persist in
an area unless they are directly or indirectly connected to a primary
producer within their associated food web (Power 1992). Thus, given that
herbivores are the main connection between plant resources (directly
limited by environmental conditions) and predators (Dobson 2009; Scott
et al. 2018), the range of a predator (omnivore or carnivore) depends on
the overlapping ranges of its herbivore preys. If sections of a
predator's range does not overlap with at least one of its prey it will
become disconnected from primary producers, and therefore we would not
expect the predator to occur in this area.

This mismatch can be the result of different mechanisms, like the
misestimation of both the predator's and the preys' ranges (Ladle and
Hortal 2013; Rondinini et al. 2006), taxonomic errors (Isaac, Mallet,
and Mace 2004; Ladle and Hortal 2013), or the lack of information about
trophic links (i.e., the lack of connection between the ranges of a
predator and a primary producer may be due a third species we don't know
is connected to both). Here in this proof of concept, we investigate the
disagreements between available data for species that compose a
well-known foodweb in the African continent, discuss the mechanisms that
can lead to this, and reinforce the importance of open geographically
explicit interaction data.

@@

We indentified areas of data deficits on the ranges of predators based
on a simple rule: we removed any part of a predator's range that did not
intersect with the range of at least one prey herbivore species, which
in turn is directly connected to a primary producer (plants). To do
that, we used a Serengeti food web dataset (Baskerville et al. 2011)
(which comprises carnivores, herbivores, and plants from Tanzania) and
its species ranges from IUCN. Then, we calculated the difference in
range sizes between the original IUCN ranges of predators and those
without the areas where they would be alone, based on species
interaction data. Finally, we added the GBIF occurrence points for the
Serengeti species to investigate whether the results would be different
if we used another source of distribution data. \#\# Data

@@

For each species in the dataset we collated point observation data from
GBIF (www.gbif.org). We used the GBIF download API to retrieve all
species occurrences on November 22nd 2022 (GBIF.org 2022). We restricted
our query to the data with spatial coordinates and which were inside the
spatial extent of our rasters. A few observations were localized in the
ocean near latitude 0° and longitude 0°. We assumed these were errors
and removed all observations falling in the extent between latitudes 2°S
and 2°N and longitudes 2°W and 2°E to keep only mainland sites. We did
not use any additional geographical filters to retrieve as much data as
possible. Being mindful of the recent and remarkable anthropogenic
impact on African megafauna, we decided to restrict the occurrences used
on the validation step to those recorded after the year 2000 (and,
therefore, only records with date information). This decision was made
after evaluating the overall temporal distribution of the GBIF records.

@@

Mammal species found in the Serengeti food web are widespread in Africa,
especially in grasslands and savannahs (panel \emph{(a)} of
fig.~\ref{fig:richness}). From our analysis, most local networks
(69.07\%) built using the original IUCN range maps had at least one
mammal species with a path to a primary producer (panel \emph{(b)} of
fig.~\ref{fig:richness}), which reinforces that the interactions we
observe in the Serengeti foodweb is representative of the interactions
for these mammals in the whole African continent. On average, local food
webs had almost half of their mammal species disconnected from basal
species (mean = 46.2\%, median = 33.3\%). In addition, 16.6\% of the
networks only had disconnected mammals, and the number of mammal species
varied from 1 to 28, with a mean of 6.7. As expected, the proportion of
carnivores with a path to a primary producer was conditional on the
total number of mammal species in each local network (third panel of
fig.~\ref{fig:richness}).

@@

\hypertarget{specialized-predators-have-higher-rates-of-range-mismatch}{%
\subsection{Specialized predators have higher rates of range
mismatch}\label{specialized-predators-have-higher-rates-of-range-mismatch}}

\begin{figure}
\hypertarget{fig:degree}{%
\centering
\includegraphics{figures/rel_loss-outdegree-species.png}
\caption{Negative relationship between the out degree of predator
species and their relative range mismatch. More specialized predators
``lose'' a higher proportion of their ranges due to mismatches with the
ranges of their preys.}\label{fig:degree}
}
\end{figure}

If we consider that we can not use areas where there are no
superposition between predators and prey on ecological analyses, we lose
more range area for predators with fewer prey (fig.~\ref{fig:degree}).
For instance, both \emph{Leptailurus serval} and \emph{Canis mesomelas}
have only one prey in the Serengeti food web (tbl.~\ref{tbl:everyone}),
each of them with a very small range compared to those of their
predator. This discrepancy between range sizes promotes significant
range loss. On the other hand, predators of the genus \emph{Panthera}
are some of the most connected species, and they also lose the least
proportion of their ranges. This mismatch between predators and preys
can also be a result of taxonomic disagreement between the geographical
and ecological data. Although \emph{Canis aureus} has the same number of
prey as \emph{Caracal caracal}, none of the prey taxa of the former
occurs inside its original range (tbl.~\ref{tbl:everyone}), which
results in complete range loss.

@@

The proportion of GBIF pixels (pixels with at least one GBIF occurrence)
matching the IUCN ranges varied a lot for species with small ranges and
way less for species with large ranges (fig.~\ref{fig:gbif}, left). This
means that species with large ranges had more area where their datasets
for ecological and geographical information agreed. The lowest
proportions of GBIF pixels occurred for species with small ranges.
Amongst herbivores, \emph{Rhabdomys pumilio} has a proportion of 25.6\%
of its presence pixels within its IUCN range, while predators have this
proportion above 47\% (such as \emph{Lycaon pictus}, with 47.6\%, and
\emph{Panthera leo}, with 49.3\%). Nevertheless, some species with
smaller ranges showed high data overlap (such as \emph{Canis mesomelas},
with 94.1\%, and many herbivores). Overall, predators and preys
displayed similar overlap variations, and species with median and large
ranges had higher proportions of agreement between GBIF, IUCN and
interaction datasets.

The proportion of GBIF pixels in revised ranges can only be equal to or
lower than that of the original ranges, as our analysis removes pixels
from the original range and does not add new ones. Rather, the absence
of a difference between the two types of ranges indicates that no pixels
with GBIF observations, hence likely true habitats, were removed by our
analysis. Here this proportion was mostly similar to that of the
original IUCN ranges for most predator species (fig.~\ref{fig:gbif}).
Two species showed no difference in proportion (\emph{Lycaon pictus} and
\emph{Panthera leo}) while four species showed only small differences
(\emph{Crocuta crocuta} lost 0.4\% of the original data overlap;
\emph{Caracal caracal} lost 3.4\%; \emph{Acinonyx jubatus} and
\emph{Panthera pardus} lost 6.2\%). On the other hand, three species,
\emph{Canis aureus}, \emph{Canis mesomelas}, and \emph{Leptailurus
serval} showed very high differences, with overlaps lowered by 100\%,
58.4\%, and 100\% respectively. These last two species are also the only
predators with a single prey in our metaweb. \emph{Canis aureus} has
four preys, but it has one of the smallest ranges in IUCN, which is not
covered by any of its preys. This result reinforces the concern raised
in the literature on the use of IUCN range maps for species that are not
well known (Herkt, Skidmore, and Fahr 2017), demonstrating how small
range species are likely to have their distribution underestimated in
the IUCN database. Additionally, the fact that \emph{Canis aureus} had
such a conspicuous discrepancy between its original IUCN range and those
of its preys, and between GBIF and IUCN data, may indicate a taxonomic
incongruency between the three databases used here, which we explore in
the Discussion section. Our results delineate how a mismatch between
GBIF and IUCN databases differ greatly with small changes in herbivore
species ranges, and it is somewhat positively related to range size for
predator species. Moreover, we show that accounting for interactions
does not necessarily aggravates this dissimilarity, but it is relevant
for species about which we have little ecological information or for
specialists groups.

\begin{figure}
\hypertarget{fig:gbif}{%
\centering
\includegraphics{figures/gbif_panels.png}
\caption{Left panel: Distribution of the proportion of GBIF pixels
(pixels with at least one occurrence in GBIF) supperposed by the IUCN
range data for different range sizes. Right panel: Differences between
the proportion of GBIF pixels matching the original and cropped IUCN
range maps for every predator species. Arrows go from the proportion
inside the original range to the proportion inside the revised range,
which can only be equal or lower. Overlapping markers indicate no
difference between the types of layers. Species markers are the same on
both figures, with predators presented in distinct colored markers and
all herbivores grouped in a single grey marker. Pixels represent a
resolution of 0.5 degrees.}\label{fig:gbif}
}
\end{figure}

@@

The case of the golden jackal (\emph{Canis aureus}) is a good
illustration of how the taxonomic, geographical and ecological data can
be used to validate one another. The jackal is a widespread taxon in
northern Africa, Europe, and Australasia, generally well adapted to
local conditions due to its largely varied diet (Tsunoda and Saito 2020;
Krofel et al. 2021). Because of that, we expected that
the~\emph{Canis}~species in our dataset would be the ones losing the
least amount of range, with a higher value of the proportion of GBIF
pixels within their IUCN range maps. However, the taxonomy of this group
is a matter of intense discussion, as molecular and morphological data
seem to disagree in the clustering of species and subspecies (Krofel et
al. 2021; Stoyanov 2020). This debate probably influenced our results:
with originally only 64.9\% of the GBIF pixels of the golden jackal
overlapping with its IUCN data, we suspect that many of the GBIF
occurrences refer to other~\emph{Canis}~species, and that its taxonomic
identification in the network database is probably outdated. This led to
a complete exclusion of~\emph{Canis aureus}~from its original range in
our analysis, despite the fact that this species has four documented
preys in our metaweb.

@@

First, \emph{Panthera leo} was one of the species with no difference
between ranges before and after our analysis, but 50.7\% of its IUCN
range map is not covered by the GBIF occurrences for this species, and
26.4\% of the GBIF point data does not overlap with its IUCN range map
(fig.~\ref{fig:gbif}). In this particular case, the IUCN maps seem to
agree with species interaction data. However, the disagreement between
the IUCN and the GBIF databases is concerning and suggests that the IUCN
maps might overestimate the lion's distribution.

On the other hand,~\emph{Leptailurus serval}~and~\emph{Canis
mesomelas}~are two of the three species that have the higher proportion
of mismatched range due to the lack of paths to a herbivore, but are
also some of the species with the higher proportion of GBIF occurrences
inside their original IUCN range maps (fig.~\ref{fig:gbif}). This
indicates that the information we are missing for these two species is
related to either an additional interaction or to the presence of
external interacting species. To illustrate that, we mapped the GBIF
data for the prey of~\emph{Leptailurus serval}, with a mobility buffer
around each point (fig.~\ref{fig:serval}). When considering GBIF data,
approximately 36\% of the prey's occurrences are within the portion of
the predator's range that was divergent from its original IUCN data.
With the buffer area, this corresponds to 5.57\% of the mismatched area.
By adding GBIF information for the prey, we could therefore reduce the
discrepancy of the range (or information) for the predator by 5.57\%
since its distribution is conditional on the occurrence of its preys. In
other words, the range mismatch was exagerated because we were missing
information on the presence of an interacting species (i.e., this also
indicates that there is a mismatch - or complementarity - between the
IUCN and GBIF data for their prey).

@@

Here we demonstrated how we can detect areas of data deficit in species
distribution data using ecological interactions. Knowing where
questionable occurrence data are can be crucial in ecological modelling
(Hortal 2008; Ladle and Hortal 2013), and accounting for these errors
can improve model outputs by diminishing the error propagation (Draper
1995). For instance, we believe our method is a way to account for
ecological interactions in habitat suitability models without making the
models more complex, but by making sure (not assuming) that the input
data - the species occurrence - actually accounts for ecological
interactions. Another application of this method is mapping areas where
data are deficient and helping to indicate priority sampling locations
for interaction data, which can, in turn, reduce uncertainty in network
prediction. For example, if a certain pixel confirms the presence of a
species both with IUCN and GBIF data, but lacks connection between
species, this pixel has a high potential to hide an unobserved
interaction and should therefore be a priority sample location.\\
It is important to notice, however, that the quality and usefulness of
this method are highly correlated with the amount and quality of data
available about species' occurrences \textbf{and} interactions. With
this paper, we hope to add to the collective effort to decode the
encrypted message that is the occurrence of a species in space and time.
A promising avenue that adds to our method is the prediction of networks
and interactions in large scales (Strydom et al. 2021; Windsor et al.
2022), for they can add valuable information about ecological
interactions where they are missing. Additionally, in order to achieve a
robust modelling framework towards actual species distribution models we
should invest in efforts to collect and combine open data on species
occurrence and interactions (Windsor et al. 2022), especially because we
may be losing ecological interactions at least as fast as we are losing
species (Valiente-Banuet et al. 2015).

@@

We acknowledge that this study was conducted on land within the
traditional unceded territory of the Saint Lawrence Iroquoian,
Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak nations. We
thank the editor and reviewers for their thoughtful comments, which
considerably improved this manuscript.
